Holomorphic field theories and higher algebra

### 2023-08-17 highervertexops.tex
not sure what i did since i last pushed

The punctured affine space $\mathring{\AA}^2$ contains a natural real hypersurface given by the unit sphere $S^3 = \{|w_1|^2 + |w_2|^2 = 1\}$, with the $w_i$ being holomorphic coordinates on $\AA^2$.
The theory of CR manifolds provides an analogue of the Dolbeault complex, 
and Zeng works with polynomial functions and polynomial forms within it:\footnote{Zeng claims to compute the CR cohomology but he in fact just works with the direct sum of the $SU(2)$-representations therein, and hence he works with this algebraic subcomplex instead.}
\beqn
\Omega^{0,\bu}_b (S^3) = \big(\C[w_i,\Bar{w}_i, \xi_i] \slash (w_i \Bar{w}_i = 1, w_i \xi_i = 0) \big) .
\eeqn
where the $\xi_i$ have degree 1 (and represent $\d \Bar{w}_i$) and where the differential is $\dbar_b = \xi_i \del_{\Bar{w}_i}$.
\owen{I'm not sure why it must provide a model for $\RR \Gamma(\mathring{\AA}^2)$, but I give some heuristic reasoning. Maybe Zeng gives a reference or proof?}
Note that one can see that polynomials in the coordinates, both holomorphic $w_i$ and antiholomorphic $\Bar{w}_i$, span the complex-valued polynomial functions on $\CC^2 \cong \RR^4$.
Their restriction to the unit-sphere $S^3$ wholly determines their behavior as functions on affine space, and hence also on punctured affine space. 
The elements $\xi_i$ correspond to $d\Bar{w}_i$ in the Dolbeault complex of punctured affine space.
The differential $\dbar_b$ then imposes the constraint, in degree zero, that these functions be holomorphic,
and its image is the image of $\dbar$ intersected with polynomial  forms in $d \Bar{w}_i$.
Thus the cohomology of this complex $\Omega^{0,\bu}_b (S^3)$ is a model for $\RR \Gamma(\mathring{\AA}^2)$.

@@

There is a more concrete way to describe this decomposition.
Consider the differential operator
\beqn
\triangle = \del_{z_1} \del_{\lambda_1} + \del_{z_2} \del_{\lambda_2}
\eeqn
acting on $\C[z_i,\lambda_i]$.
\owen{This isn't the image of the quadratic Casimir. In what sense is it the relevant Laplacian? On the other hand, it is a weight zero operator, even if it lowers the bigrading by (1,1).}
Let $H_{a,b} \subset V_{a,b}$ denote the harmonic polynomials of bidegree $(a,b)$.
Examples of harmonic polynomials include $z_1^a \lambda_2^b$ for arbitrary $a,b$ and
$z_1 \lambda_1 - z_2 \lambda_2$ for $a = 1 = b$.
A direct computation shows that there is a direct sum decomposition \owen{of representations?}
\begin{align*}
V_{a,b} & = H_{a,b} \oplus |(z,\lambda)|^2 V_{a-1,b-1} \\
& = H_{a,b} \oplus |(z,\lambda)|^2 H_{a-1,b-1} \oplus |(z,\lambda)|^4 H_{a-2,b-2} \oplus \cdots 
\end{align*}
(Compare with the well-known construction of spherical harmonics.)

Consider now  
\[
|(z,\lambda)|^2 \define z_1 \lambda_1 + z_2 \lambda_2, 
\]
which describes the evaluation pairing. 
It is preserved by the $SL(2)$-action, and hence we see that the canonical quotient map 
\[
q: \C[z_i,\lambda_i] \to R = \C[z_i,\lambda_i]/(|(z,\lambda)|^2 = 1)
\]
is $SL(2)$-equivariant.
Moreover, inside $V_{a,b}$, any multiple of $|(z,\lambda)|^2$ is annihilated by $q$,
so $R$ is spanned by the images of $H_{a,b}$.
Thus, there is a direct sum decomposition $\sfR = \bigoplus_{a,b} \sfR_{a,b}$ where $\sfR_{a,b} = q(H_{a,b})$ is the image of the subspace of harmonic polynomials in $\C[z_i,\lambda_i]$ under the quotient map.

@@

Let $\sfH \define H^\bu(\sfA)$ and let $i,p$ be the $SU(2)$-equivariant inclusions and projections $i \colon \sfH_2 \hookrightarrow \sfA$, $p \colon \sfA \to \sfH$, given by the direct sum description of the corollary above.

@@

The model $\sfA$ for algebraic functions on punctured affine space was constructed using the ring
\beqn
\sfR = \C[z_i,\lambda_i] \slash (z_i \lambda_i = 1) .
\eeqn
Let $\Hat{\sfR}$ denote the completion of this ring with respect to the ideal generated by $z_1,z_2$.
So I think
\beqn
\Hat{\sfR} = \C[[z_i]][\lambda_i] \slash (z_i \lambda_i) .
\eeqn

@@

There is a natural deformation quantization of $\cP$ as a Weyl algebra $\cW$.
Observe that there is a dg Lie subalgebra
\[
\cW_2 = (\Sym^2(V \oplus V^*) \oplus \Sym^0(V \oplus V^*), [-,-])
\]
of $\cW$.
It is a central extension of $\cP_2$, and let $\omega \in H^2(\cP_2)$ denote the associated 2-cocycle.

@@

\begin{lem}
Given a Lie algebra map $J: \fg \to \cP_2$, the obstruction to lifting it to a map of Lie algebras $\fg \to \cW_2$ is the 2-cocycle $J^* \omega \in H^2(\fg)$.
\end{lem}

>>>>

The punctured affine space $\mathring{\AA}^2$ contains a natural real hypersurface given by the unit sphere $S^3 = \{|w_1|^2 + |w_2|^2 = 1\}$, with the $w_i$ being holomorphic coordinates on $\AA^2$.
The theory of CR manifolds provides an analogue of the Dolbeault complex, 
and Zeng works with polynomial functions and polynomial forms within it:\footnote{Zeng claims to compute the CR cohomology but he in fact just works with the direct sum of the $SU(2)$-representations therein, and hence he works with this algebraic subcomplex instead.}
\beqn
\Omega^{0,\bu}_b (S^3) = \big(\C[w_i,\Bar{w}_i, \xi_i] \slash (w_i \Bar{w}_i = 1, w_i \xi_i = 0) \big) .
\eeqn
where the $\xi_i$ have degree 1 (and represent $\d \Bar{w}_i$) and where the differential is $\dbar_b = \xi_i \del_{\Bar{w}_i}$.
Note that one can see that polynomials in the coordinates, both holomorphic $w_i$ and antiholomorphic $\Bar{w}_i$, span the complex-valued polynomial functions on $\CC^2 \cong \RR^4$, i.e., as a real vector space.
Their restriction to the unit-sphere $S^3$ wholly determines their behavior as functions on $\RR^4$, and hence also on punctured affine space. 
The elements $\xi_i$ correspond to $d\Bar{w}_i$ in the Dolbeault complex of punctured affine space.
The differential $\dbar_b$ then imposes the constraint, in degree zero, that these functions be holomorphic,
and its image is the image of $\dbar$ intersected with polynomial  forms in $d \Bar{w}_i$.
Thus the cohomology of this complex $\Omega^{0,\bu}_b (S^3)$ is a model for $\RR \Gamma(\mathring{\AA}^2)$, by a result of~\cite{FHK}.

@@

There is a more concrete way to describe this decomposition, 
modeled on the construction of spherical harmonics as representations of $SO(n)$.
Consider the differential operator
\beqn
\triangle = \del_{z_1} \del_{\lambda_1} + \del_{z_2} \del_{\lambda_2}
\eeqn
acting on $\C[z_i,\lambda_i]$;
it is inspired by the Laplacian on $\CC^2$ when we view $\lambda_i$ as a version of $\Bar{z_i}$.
This operator has weight zero and, in fact, it is $SL(2)$-equivariant.
\owen{I need to check this myself. If there's a slick way, we should include it.}
On the other hand, it maps $V_{a,b}$ to $V_{a-1,b-1}$, i.e., it changes the bigrading by~$(-1,-1)$.

Let $H_{a,b} = \ker \triangle|_{V_{a,b}}$ denote the ``harmonic'' polynomials of bidegree $(a,b)$.
Examples of harmonic polynomials include $z_1^a \lambda_2^b$ for arbitrary $a,b$ and
$z_1 \lambda_1 - z_2 \lambda_2$ for $a = 1 = b$.
A direct computation shows that there is a direct sum decomposition \owen{of representations?}
\begin{align*}
V_{a,b} & = H_{a,b} \oplus |(z,\lambda)|^2 V_{a-1,b-1} \\
& = H_{a,b} \oplus |(z,\lambda)|^2 H_{a-1,b-1} \oplus |(z,\lambda)|^4 H_{a-2,b-2} \oplus \cdots 
\end{align*}
(Compare with the well-known )

Consider now  
\[
|(z,\lambda)|^2 \define z_1 \lambda_1 + z_2 \lambda_2, 
\]
which describes the evaluation pairing. 
It is preserved by the $SL(2)$-action, and hence we see that the canonical quotient map 
\[
q: \C[z_i,\lambda_i] \to \sfR = \C[z_i,\lambda_i]/(|(z,\lambda)|^2 = 1)
\]
is $SL(2)$-equivariant.
Moreover, inside $V_{a,b}$, any multiple of $|(z,\lambda)|^2$ is annihilated by $q$,
so $R$ is spanned by the images of $H_{a,b}$.
Thus, there is a direct sum decomposition $\sfR = \bigoplus_{a,b} \sfR_{a,b}$ where $\sfR_{a,b} = q(H_{a,b})$ is the image of the subspace of harmonic polynomials in $\C[z_i,\lambda_i]$ under the quotient map.

@@

Let $\sfH \define H^\bu(\sfA)$ and let $i,p$ be the $SU(2)$-equivariant inclusions and projections $i \colon \sfH \hookrightarrow \sfA$, $p \colon \sfA \to \sfH$, given by the direct sum description of the corollary above.

@@

The model $\sfA$ for algebraic functions on punctured affine space was constructed using the ring
\beqn
\sfR = \C[z_i,\lambda_i] \slash (\sum_i z_i \lambda_i = 1) .
\eeqn
Let $\Hat{\sfR}$ denote the completion of this ring with respect to the ideal generated by $z_1,z_2$.
So I think
\beqn
\Hat{\sfR} = \C[[z_i]][\lambda_i] \slash (\sum_i z_i \lambda_i) .
\eeqn

@@

There is a natural deformation quantization of $\cP$ as a Weyl algebra $\cW$.
Observe that there is a dg Lie subalgebra
\[
\cW_2 = (\Sym^2(V \oplus V^*) \oplus \Sym^0(V \oplus V^*), [-,-])
\]
of $\cW$.
It is a central extension of $\cP_2$; let $\omega \in H^2(\cP_2)$ denote the associated 2-cocycle.

@@

\begin{lem}
Given a Lie algebra map $J: \fg \to \cP_2$, the obstruction to $J$ being a map of Lie algebras $\fg \to \cW_2$ is the 2-cocycle $J^* \omega \in H^2(\fg)$.
\end{lem}
